---
title: "Oral Health Project"
author: "Rasha Elnimeiry"
date: "December 2019"
output:
    flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---



```{r setup, include=FALSE}
library(flexdashboard)
library(leaflet)
library(tidyverse)
library(ggplot2)
library(rgdal)
library("raster")
library("sp")
library("GISTools")
library("raster")
library("sp")
library("GISTools")
library(tidyverse)
library(sf)
library(units)
library(tmap)

setwd("/Users/rashaelnimeiry/Desktop/Meetings ND/ND county DD NDGISHubData")
# cnty<-readOGR("NDHUB.COUNTIES_polygon.shp")
setwd("/Users/rashaelnimeiry/Desktop/Meetings ND")
cnty<-readOGR("cntyshpschlsUnique.shp")
#has counts of dental offices and schools dental sealant prog counts
daddn<-read.csv("daddn.csv")
daddn<-subset(daddn,  officestate=="ND")
allschools<-read.csv("geoallschools.csv")

# 
# cntyna <- c("Burke"	, "Billings"	,"Bowman"	,"Ward"	,"Rolette"	,"Mountrail"	,"Richland"	,"Towner"	,
#             "Ramsey"	,"Stutsman"	,"Oliver"	,"Sioux"	,"Adams"	,"Steele"	,"Barnes"	,"Hettinger"	,
#             "Ransom"	,"Pembina"	,"Dickey"	,"Wells"	,"Benson"	,"Grand Forks"	,"McIntosh"	,
#             "Logan"	,"Renville"	,"Pierce"	,"Foster"	,"Grant"	,"Golden Valley"	,"Divide"	,"Stark"	,
#             "Nelson"	,"Sargent"	,"LaMoure"	,"Bottineau"	,"Cavalier"	,"Sheridan"	,"McHenry"	,
#             "Traill"	,"Eddy"	,"Emmons"	,"Dunn"	,"Walsh"	,"Kidder"	,"Cass"	,"Mercer"	,"McLean"	,
#             "McKenzie"	,"Burleigh"	,"Morton"	,"Slope"	,"Williams"	,"Griggs"	)
# cnty@data<-cbind(cntyna, cnty@data)

icon<- makeIcon(
  iconUrl = "https://cdn2.iconfinder.com/data/icons/human-body-parts/614/5008_-_Tooth-512.png",
  iconWidth = 22, iconHeight = 21,
  iconAnchorX = 0, iconAnchorY = 0)


pal<- colorNumeric(palette = c("#087482", "#796e66", "#faa21b", "#d34727", "#a8353a"), domain = cnty@data$POP2010, na.color="transparent")

mytext<-paste(
  "<b>","County Name: ","</b>", cnty@data$cntyna,"<br/>",
  "<b>","Population: ","</b>", round(cnty@data$POP2010, 2), 
  sep = "") %>% 
  lapply(htmltools::HTML)


org <- "#d34727"
dgrey <- "#796e66"
lgrey <- "#b6b0a2"
turq <- "#087482"
blk <- "#000000"
burg <- "#a8353a"
byell <- "#faa21b"
bblue <- "#049fda"
dgrn <- "#709749"
lgrn <- "#b3bd35"

```
Dental Offices
=====================================
Column {data-width=650}
-----------------------------------------------------------------------

### North Dakota Dental Office Locations

```{r}


leaflet(cnty) %>%
  setView(lng= -100.400, lat=47.6715, zoom=7.5) %>% 
   addPolygons(data=cnty, color= dgrey,  weight = 2, fillColor = org, fillOpacity = .1, opacity = .30, 
               highlightOptions = highlightOptions(color = turq, weight = 3.5, opacity = .6,
                 bringToFront = F), 
               label = mytext,
               labelOptions = labelOptions( 
                 style = list("font-weight" = "normal", padding = "3px 8px"), 
                 textsize = "13px", 
                 direction = "auto" )) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers( icon= icon,
    lng = daddn$lon,
    lat = daddn$lat,
    popup = paste("<b>", "Office Name: ","</b>", daddn$officename , "<br>" ,
                  "<b>", "Specialty: ", "</b>", daddn$specialtyada, "<br>",
                  "<b>", "Practicing Status: ", "</b>", daddn$Practicing,  "<br>",
                  "<b>", "Licensure: ", "</b>", daddn$license), 
    group="Dental Office Locations")      %>%     

addPolygons( data=cnty ,
             fillColor = ~pal(cnty@data$POP2010),
             weight=0.9,
             stroke=TRUE, 
             color=dgrey,
             fillOpacity = 0.5, 
             label = mytext,
             labelOptions = labelOptions( 
               style = list("font-weight" = "normal", padding = "3px 8px"), 
               textsize = "13px", 
               direction = "auto" ),
             highlightOptions = highlightOptions(color = "white", weight = 4.5, opacity = .6,
                                                 bringToFront = T),

  group="Population 2010")    %>% 
  
  addLegend("topright", 
            pal = pal,
            values=cnty@data$POP2010,
            labels = ~pal, 
            title="North Dakota Population 2010",
            opacity = .5,
  group="Legend") %>% 

addLayersControl(overlayGroups = c( "Dental Office Locations", "Population 2010", "Legend")) 

```

Column {data-width=150}
-----------------------------------------------------------------------

### County Population

```{r}
DT::datatable(cnty@data[, c("NAME", "cntdntf", "POP2010", "POP10_S") ],
  rownames = F, colnames= c("County" ,"Dental Offices per County", "Population", "Population per Sq. mi."), options = list(pageLength = 53)
)
```

### Notes

```{r}
"Dental Office Locations in North Dakota"
```

<b>How to Use:</b> 
</br>
</br>* Click on the "layers" icon to toggle on/off different layers on the chloropleth map. 
</br>
</br>* Click the "tooth" icon to display more information about the dental office at that location.
</br>
</br>* Zoom in and out using "+" or "-" at the top left of map, or pressing and holding the "CNTRL" key and scrolling the wheel on your mouse up or down.
</br>
</br><b>Data Sources:</b> 
</br>
</br>* Population counts are per U.S. Census, 2010.
</br>
</br>* Dental office "locations" and "counts" are as per ND Department of Health, October, 2019.
</br>
</br>
</br>Updated on December 8, 2019.

-----------------------------------------------------------------------

School-Based Dental Sealant Program
=====================================
Column {data-width=650}
-----------------------------------------------------------------------

### North Dakota School-Based Dental Sealant Program

```{r}


library(htmltools)
library(ggmap)
library(leaflet)

pal<- colorNumeric(palette = c("#087482", "#796e66", "#faa21b", "#d34727", "#a8353a"), domain = cnty@data$POP2010, na.color="transparent")

pal2<-colorFactor(palette =c(blk, turq, org, byell, "white"), domain=levels(allschools$Provider.Type))


cntyna <- c("Burke"	, "Billings"	,"Bowman"	,"Ward"	,"Rolette"	,"Mountrail"	,"Richland"	,"Towner"	,
            "Ramsey"	,"Stutsman"	,"Oliver"	,"Sioux"	,"Adams"	,"Steele"	,"Barnes"	,"Hettinger"	,
            "Ransom"	,"Pembina"	,"Dickey"	,"Wells"	,"Benson"	,"Grand Forks"	,"McIntosh"	,
            "Logan"	,"Renville"	,"Pierce"	,"Foster"	,"Grant"	,"Golden Valley"	,"Divide"	,"Stark"	,
            "Nelson"	,"Sargent"	,"LaMoure"	,"Bottineau"	,"Cavalier"	,"Sheridan"	,"McHenry"	,
            "Traill"	,"Eddy"	,"Emmons"	,"Dunn"	,"Walsh"	,"Kidder"	,"Cass"	,"Mercer"	,"McLean"	,
            "McKenzie"	,"Burleigh"	,"Morton"	,"Slope"	,"Williams"	,"Griggs"	)
cnty@data<-cbind(cntyna, cnty@data)


mytext<-paste(
 "<b>", "County Name: ","</b>", cnty@data$cntyna,"<br/>",
 "<b>","Population: ", "</b>", round(cnty@data$POP2010, 2), 
  sep = "") %>% 
  lapply(htmltools::HTML)

leaflet(cnty) %>%
  setView(lng= -100.400, lat=47.6715, zoom=7.5) %>% 
  addPolygons(data=cnty, 
              fillColor = "grey",
               weight=0.9,
               smoothFactor = 1, 
               color="grey",
               fillOpacity = 0.5,
              highlightOptions = highlightOptions(color = turq, weight = 3.5, opacity = .6,
                 bringToFront = F), 
               label = mytext,
               labelOptions = labelOptions( 
                 style = list("font-weight" = "normal", padding = "3px 8px"), 
                 textsize = "13px", 
                 direction = "auto" )) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
          
  addCircleMarkers(
    radius=8.5,
    weight=.9,
    fillOpacity = .42,
    lng = allschools$lon,
    lat = allschools$lat,
    color= ~pal2(allschools$Provider.Type),
    label= lapply(paste("<b>","School Name: ", "</b>", allschools$Name.of.School , "<br>" ,
                   "<b>","Provider: ","</b>", allschools$Provider.Type, "<br>"), HTML)  , 

 group="School Locations") %>% 
  addLegend("topright", 
            pal = pal2,
            values=allschools$Provider.Type,
            labels = ~pal2, 
            title="North Dakota Schools Enrolled in Dental Sealant Programs by Provider Type",
            group="Legend",
            opacity = .5) 

```

Column {data-width=150}
-----------------------------------------------------------------------

### County Population

```{r}
DT::datatable(cnty@data[, c("cntyna", "cntschl", "POP2010", "POP10_S")],
  rownames = FALSE, colnames=c("County","Schools Enrolled in Dental Sealant Program Per County", "Population", "Population per Sq. mi." ), options = list(pageLength = 53)
)
```


### Notes

```{r}
 "School Locations and Dental Sealant Program Providers in North Dakota"

```

<b>How to Use:</b> 
</br>
</br>
* Click the "circle" icon to display more information about the school and the type of dental sealant program provider at that school location. 
</br>
</br>* Zoom in and out using "+" or "-" at the top left of map, or pressing and holding the "CNTRL" key and scrolling the wheel on your mouse up or down.
</br>
</br><b>Data Sources:</b> 
</br>
</br>* Population counts are per U.S. Census, 2010.
</br>
</br>* Dental office "locations" and "counts" are as per ND Department of Health, October, 2019.
</br>
</br>
</br>Updated on December 6, 2019.

